### 2020-07-24 documents/pamphlet.tex
Eppenwolf milled out

\documentclass[fleqn,10pt]{paper}

\usepackage[left=2cm,right=2cm,
top=0cm,
bottom=2.25cm,%
headheight=11pt,%
letterpaper]{geometry}

\include{preamble}

\title{On pulsed microwave acoustic resonant viral inactivation: summary}

\begin{document}

\maketitle

%\footnote{The}

This is a heavily condensed version of a (hideously unfinished) paper at \href{https://www.github.com/0xDBFB7/covidinator/documents/paper.pdf}{\faGithub/0xDBFB7/covidinator}.\footnote{Daniel Correia [TODO affiliation] \orcidlink{0000-0002-9353-0216}}

Both the microwave and dosimetric components \footnote{field field?} appear to be unusually sensitive to experimental design and exceptionally prone to artifacts that cause false positive results: \cite{Microwave1982} $\neq$ \cite{Resonances1987}, \cite{Effects1985a} $\neq$ \cite{Cytogenetic1986}, \cite{Comprehensive2018}. Though the cited articles have taken care to avoid these effects, is possible that none of what is discussed below actually exists.

{\Large \textbf{The mechanism}}

The literature is littered with irreproducible claims of non-thermal effects; however, this appears to be a singular anomaly.

\cite{Microwave2009} \textrightarrow \ (\cite{focusing2014} $\parallel$ \cite{Efficient2015} $\parallel$ \cite{Resonant2017})\footnote{A great many papers are involved.} appear to establish that the net charge, mass distribution, and envelope stiffness of a few species of viruses supports a weak resonance mode in the microwave spectrum, with a peak at roughly 8 GHz.

They appear to demonstrate that this allows a comparatively inconsequential electromagnetic field magnitude to produce mechanical stresses sufficient to crack the lipid envelope, destroying the virus.

\cite{Efficient2015} specifically provides a reasonable spherical-cow model for how this could occur that agrees well with experiment.

This work can be extended with the following assertion: the virus' Q factor is only 2, meaning that the steady-state amplitude should be reached in some nanoseconds.

Barring protein mechanical fatigue [] or other mechanisms that could prolong the inactivation duration, pulsing the field at nanosecond scales, it appears to reduce the threshold to less than 1/Xth of all safety limits. Better yet, by scanning the focal point, a significant area can be treated with a relatively small transmit power.

\cite{Efficient2015} test with various strains of Influenza A, we test with Coliphage T4; structural similarities with NCoV mean this is may be transferable.

These authors above deserve all the credit for this finding; nothing in this work is even remotely novel.

{\Large \textbf{How is this better than the alternatives?}}

It produces no ozone and causes no damage to tissue (unlike near UV) even during continuous exposure, penetrates tissue (unlike far UV and cold-plasma), and

{\Large \textbf{Is it safe?}}

The consensus is that the environment not favorable to create such modes; damping from the surrounding fluids is too strong \cite{Vibrational2002}.

Quality in-vitro data supports this assertion. \cite{Cytogenetic2006} expose blood cells (a representative cell) to pulse power in precisely the regime required in this work; 8.2 GHz, 8 ns duration, 50 khz repetition rate, a whopping pulse power density of 250,000\Wsqm\footnote{computed from average power / duty cycle} (2500x the time-averaged power density safety limit), average power density 100\Wsqm (the safety limit), for 2 hours, finding no change in any of the measured quantities.

Similarly, \cite{DNA2004} use 8.8 GHz, 180 ns pulse width, peak power 65,000 W, repetition rate 50 Hz, exposure duration 40 min, and find genotoxicity purely due through the expected thermal mode.

[Foster 1987]

\footnote{These are cherry-picked.}

The literature reviews of \cite{ICNIRP2020} and \cite{C95} that they have conducted are far more exhaustive than ours.

Whether this is due to a specific feature of the virion or simply, we are unsure. The lipid bilayer of the virion differs in composition from that of the host cell.

This works because the damage mechanism

{\Large \textbf{How do we wield this sword?}}

If this is correct, any region with a power density in the 7 to 9 GHz band greater than 300\Wsqm \ for 10 ns will be sterilized.

This is not a particularly demanding requirement.

We provide a reference design that costs \$3 in prototype quantities.

With modern semiconductor processes, it is trivial to produce

It is possible to create these conditions within human tissues; attenuation limits the safe treatment depth to about 3 cm.

\begin{autem}
	a
\end{autem}

{\Large \textbf{Tasks}}

We can see the following

\begin{itemize}
	\item To prevent duplication of work, are any other teams working on this now? It seems a rather obvious technique. A thorough review of arxiv etc should be performed.
	\item Double-check all figures and assumptions.
	\item Get permission to use York University as affiliation, post on arxiv.
	\item Send off spectrometer (or, ideally, a more precise piece of equipment) to a BSL with a sample of NCoV to obtain the spectra and thresholds.
	\item Determine the ideal application (requires detailed sims), finish the design for the electromagnetic mask, start bronchoscopic trials?

	\item Check with the FCC/Industry Canada if we can get a temporary exemption for these intentional radiators?

\end{itemize}

%\centering {\color{red} \faHeart}

\clearpage

\printbibliography[heading=none, title={}]

\end{document}