### 2020-08-01 firmware/eppenwolf/protocols/microwave_spectrum.py_nw_.tex
Noweb

\documentclass{article}\usepackage{noweb}\pagestyle{noweb}\noweboptions{}\begin{document}\nwfilename{microwave_spectrum.py}\nwbegindocs{0}\nwdocspar% ===> this file was generated automatically by noweave --- better not edit it
\nwenddocs{}\nwbegindocs{1}\nwdocspar
Our aims with this

\nwenddocs{}\nwbegindocs{2}\nwdocspar
Inspired by \cite{Biocoder2010}, we write the microbiology protocol directly into the Eppenwolf firmware.

    log_serial.println(String(__FILE__));

}

\nwenddocs{}\nwbegincode{3}\sublabel{NW3GNx30-3GNx30-1}\nwmargintag{{\nwtagstyle{}\subpageref{NW3GNx30-3GNx30-1}}}\moddef{microwave_spectrum.py~{\nwtagstyle{}\subpageref{NW3GNx30-3GNx30-1}}}\endmoddef\nwstartdeflinemarkup\nwenddeflinemarkup
alalsdjkf

BASE_BIAS_GAINaa
aa
\nwnotused{microwave_spectrum.py}\nwendcode{}

\nwixlogsorted{c}{{microwave_spectrum.py}{NW3GNx30-3GNx30-1}{\nwixd{NW3GNx30-3GNx30-1}}}%
\end{document}