### 2020-08-01 firmware/eppenwolf/protocols/pulse_1.nw.tex

\documentclass[paper.tex]{subfiles}% ===> this file was generated automatically by noweave --- better not edit it
\begin{document}

import device_comms
import sweep

\nwfilename{pulse_1.nw}\nwbegindocs{1}\nwdocspar
Our aims with this experiment are to perform the following:

Avoid any ambiguity regarding the temperature rise seen in the

\nwenddocs{}\nwbegindocs{2}\nwdocspar
Inspired by \cite{Biocoder2010}, we write the microbiology protocol directly into the Eppenwolf firmware using literate programming.

}

\nwenddocs{}\nwbegincode{3}\moddef{microwave_spectrum.py}\endmoddef\nwstartdeflinemarkup\nwenddeflinemarkup
alalsdjkf

test()
BASE_BIAS_GAINaa
aa

\nwendcode{}\nwbegindocs{4}aaaaaaaaaaaa

aaaaaaaaaaaaa

# T - 5 hours

# Phage+Bacteria / Bacteria-only / phage only?
# expose (parameters) / sham / thermal positives

# For each of the animal and human published studies, four
# predetermined and specific quality control measures were used to
# assess the quality of the publication.
# 1. Whether or not the researchers have used ‘‘blind’’ evaluations to
# avoid individual ‘‘bias’’, B or nB;
# 2. Whether or not there was detailed/adequate description of
# ‘‘dosimetry’’ in the publication for independent replication/
# confirmation studies, D or nD.
# 3. Whether or not ‘‘positive controls’’ were included in the
# experiment(s) to replicate/confirm the observations, P or nP.
# 4. Whether or not ‘‘sham-exposed controls’’ (more appropriate to
# compare with the observations made in RF exposure conditions)
# were included in the experiments, S or nS. However, it was not our
# intention to ‘‘rank’’ the publications either to exclude or include
# the reported data in the meta-analysis.

def initialize():
    # Name this run
    # Fluidic plate ID

    print("Fill phage into all selected cuvettes.")
    Simulataneously and separately

#centrifuge down bacteria lightly to concentrate
#Mix with fresh broth
#Vortex for X seconds?
#
# Rajnovic 2019
# Overnight cultures of E. coli were centrifuged and the pellets resuspended in 0.1 mM PB to
# achieve a concentration of 10^10 cfu/mL.
# The resulting suspensions were subject to serial dilution
#
# For each assay, 160 μL of LB were mixed with
# 20 μL phage solution, 20 μL of bacteria solution and 20μL of PB in

def bacteria_alone_control():
    //repeated twice

    # fill_phage()

def bacteria_and_phage_control():
    //repeated twice

def bacteria_and_inactivated_phage_positive_control():
    //

"Blank" repeatablity test with just water in all cuvettes

Add a small amount of growing e.coli to a new nutrient broth,
aerate vigorously, then add to culture chamber.

Expose phage.
Fill culture.
Monitor culture for log stage.
Mix culture and phage.
Monitor mixture.

Types of exposure:

RF

Expose phage, mix with log bacteria, monitor.

Bare

Sham

?

Positive control

Bring phage cuvette temperature to 120c with trace.

Autoclaved

\end{document}
\nwenddocs{}