### 2020-08-01 firmware/eppenwolf/protocols/pulse_1.nw.tex
Noweb -> pweave

\documentclass[paper.tex]{subfiles}% ===> this file was generated automatically by noweave --- better not edit it
\begin{document}

\clearpage
{\Large \it Experiment 2:}\\

\nwfilename{pulse_1.nw}\nwbegindocs{1}\nwdocspar
Our aims with this experiment are to perform the following:

\begin{itemize}
\item Crudely test the pulse hypothesis. \cite{Efficient2015}'s use of a 15 minute exposure caused us some consternation.
    While T4 is quite unlike enveloped viruses, it at least provides a first approximation.
\item Avoid any ambiguity regarding the temperature rise seen in the previous paper.
\end{itemize}

In the end, performing this was a mistake; but by the time we realized this we had already put in the bulk of the time.

Inspired by \cite{Biocoder2010}, we write the microbiology protocol directly into the {\it Eppenwolf} firmware using \cite{Noweb}.

<<pulse_1.py>>
import device_comms
import sweep

\nwenddocs{}\nwbegindocs{2}\nwdocspar

# T - 5 hours

# Phage+Bacteria / Bacteria-only / phage only?
# expose (parameters) / sham / thermal positives

For each of the animal and human published studies, four
predetermined and specific quality control measures were used to
assess the quality of the publication.
1. Whether or not the researchers have used ‘‘blind’’ evaluations to
avoid individual ‘‘bias’’, B or nB;
2. Whether or not there was detailed/adequate description of
‘‘dosimetry’’ in the publication for independent replication/
confirmation studies, D or nD.
3. Whether or not ‘‘positive controls’’ were included in the
experiment(s) to replicate/confirm the observations, P or nP.
4. Whether or not ‘‘sham-exposed controls’’ (more appropriate to
compare with the observations made in RF exposure conditions)
were included in the experiments, S or nS. However, it was not our
intention to ‘‘rank’’ the publications either to exclude or include
the reported data in the meta-analysis.

def initialize():
    # Name this run
    # Fluidic plate ID

    print("Fill phage into all selected cuvettes.")
    Simulataneously and separately

#centrifuge down bacteria lightly to concentrate
#Mix with fresh broth
#Vortex for X seconds?
#
# Rajnovic 2019
# Overnight cultures of E. coli were centrifuged and the pellets resuspended in 0.1 mM PB to
# achieve a concentration of 10^10 cfu/mL.
# The resulting suspensions were subject to serial dilution
#
# For each assay, 160 μL of LB were mixed with
# 20 μL phage solution, 20 μL of bacteria solution and 20μL of PB in

def bacteria_alone_control():
    //repeated twice

    # fill_phage()

def bacteria_and_phage_control():
    //repeated twice

def bacteria_and_inactivated_phage_positive_control():
    //

"Blank" repeatablity test with just water in all cuvettes

Add a small amount of growing e.coli to a new nutrient broth,
aerate vigorously, then add to culture chamber.

Expose phage.
Fill culture.
Monitor culture for log stage.
Mix culture and phage.
Monitor mixture.

Types of exposure:

RF

Expose phage, mix with log bacteria, monitor.

Bare

Sham

?

Positive control

Bring phage cuvette temperature to 120c with trace.

Autoclaved

\end{document}
\nwenddocs{}