### 2020-08-01 documents/microwave_spectrum.tex
SA test

\documentclass[paper.tex]{subfiles}% ===> this file was generated automatically by noweave --- better not edit it
\begin{document}

\clearpage
{\Large Experiment 1:}\\

\nwfilename{microwave_spectrum.pnw}\nwbegincode{1}\moddef{microwave_spectrum.py}\endmoddef\nwstartdeflinemarkup\nwenddeflinemarkup
import device_comms
import sweep
\nwendcode{}\nwbegindocs{2}\nwdocspar

\nwenddocs{}\nwbegincode{3}\moddef{}\endmoddef\nwstartdeflinemarkup\nwenddeflinemarkup

\nwendcode{}\nwbegindocs{4}\nwdocspar

\end{document}
\nwenddocs{}