### 2020-08-12 documents/microwave_spectrum.tex
Writing the protocol

\documentclass[paper.tex]{subfiles}% ===> this file was generated automatically by noweave --- better not edit it
\begin{document}

\clearpage
{\Large Experiment 1:}\\

\nwfilename{microwave_spectrum.pnw}\nwbegincode{1}\moddef{microwave_spectrum.py}\endmoddef\nwstartdeflinemarkup\nwenddeflinemarkup
import sys
sys.path.append('/home/arthurdent/covidinator/firmware/eppendoofus/host/')

import device_comms
import sweep

\nwendcode{}\nwbegindocs{2}\nwdocspar

It cannot be overstated how sensitive this measurement is. At 12 GHz, merely bringing one's hand near the coaxial line almost doubled
the power sensor reading.

\nwenddocs{}\nwbegincode{3}\moddef{}\endmoddef\nwstartdeflinemarkup\nwenddeflinemarkup
import random

\nwendcode{}\nwbegindocs{4}\nwdocspar

We randomize \cite{first2000}

\nwenddocs{}\nwbegincode{5}\moddef{}\plusendmoddef\nwstartdeflinemarkup\nwenddeflinemarkup

\nwendcode{}\nwbegindocs{6}\nwdocspar

\end{document}
\nwenddocs{}