### 2020-08-30 documents/test.tex

\documentclass[fleqn,10pt]{paper}

\usepackage[left=2cm,right=2cm,
top=0.5cm,
bottom=1.5cm,%
headheight=11pt,%
letterpaper]{geometry}

\include{preamble}

\title{{\it Maxwell's Silver Hammer:}\\ On pulsed microwave acoustic resonant viral inactivation}

\begin{document}

\section{Silicon carbide test}

To verify that we can detect a resonance with this setup, we use the microwave susceptor mixture described by \cite{Effect2016}, which can be tuned to the correct range by varying the concentration of SiC to binder. We use the S-1 mixture.

0.6g 2000-mesh silicon carbide powder was added to 1.4 g white Elmer's glue, mixed manually until homogeneous,
applied to a mylar film, and dried with a hair dryer. A small wafer of the hardened mixture, perhaps 0.3 mm thick, 5 mm x 2 mm, was placed directly on top of the coplanar waveguide.

The spectrum was then captured, averaging (?10) sweeps of the VCO. The background spectrum was taken without the wafer in place.

\begin{figure}[H]
	\includesvg[width=\textwidth]{sic_9_1}
	\caption{}
\end{figure}

\begin{figure}[H]
	\includesvg[width=\textwidth]{sic_9_2}
	\caption{}
\end{figure}

\cite{Effect2016} measures a peak at about 7.2 GHz. Our peak is at approximately 7.849 GHz, which is well within variation in powder grain sizes and concentration.

The paper specifies the reflection loss, S$_{11}$. Unfortunately, having no directional couplers, it is probably not possible to directly determine S$_{21}$ and S$_{11}$ with this setup.

Naively, we might expect the peak to appear as a decrease at both sensors; instead it appears as an increase in transmission and decrease in reflection. It could be possible that the presence of the dielectric wafer has merely detuned an existing resonance, that the peak is a coincidence, and that these data mean nothing.
The raw voltage plots do not suggest that this is the case; but it is nonetheless concerning.

\footnote{The paper mentions that the "blending fraction of silicon carbide powders to epoxy resin was 30 \%, 35 \%, 40 \%, 45 \% and 50\% by weight". That seems a little ambiguous as to whether that's "weight per total mass" or "weight per epoxy".}

\end{document}