### 2020-09-21 documents/pseudo_soliton.tex
text

%!TeX root = pseudo_soliton
\documentclass[paper.tex]{subfiles}
\begin{document}

Precursor waves do not appear to be explicitly mentioned in current standards, but are implicitly governed by the Fourier-spectrum clauses. Ultra-short wavepackets of a similar nature are widely used in MRI systems with a repetition rate up to 300 MHz. We can make some mechanistic claims.

At 9 GHz, assuming that a torso is about 10 penetration depths in radius, implementing this would decrease the required surface field for complete eradication of the virus from the body to only 1 kV/m, rather than 6.6 MV/m ($1/\sqrt{10}=0.316$ rather than $e^{-10}=4.5e-5$).

\end{document}