### 2020-11-04 documents/fluorescence.tex
pmt counter

\begin{document}

Low background is required.

Xu's method \cite{Quantification2020}

Nanodrop, using 280 nm absorbance.

the input edge is considered a clock; the minimum pulse width limits apply, but are not clearly specified in the datasheet. In this case (1/155 mhz) = 6 ns.