### 2020-11-05 documents/acknowledgement.tex
test

%!TeX root = acknowledgement
\documentclass[paper.tex]{subfiles}
\begin{document}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Acknowledgements}\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This paper is not novel enough to deserve this chapter, but those to be acknowledged are.

The authors of the original papers deserve the credit for this entire work; it is rare that such a monumental finding is served on a silver platter.

Chiefly among the enablers are of course my parents, Doris and Brian, without whom this work would have been done by someone else, and who also supported this work financially.

Thanks to (Dr. ) P. Barfuss for spending time reviewing an early draft.

This paper would have been absolutely impossible without the support of thousands of open-source software packages, ranging from monolithic state-supported tools like Paraview to poorly-compensated but equally invaluable scripts like zonca. A special thanks to the dedicated authors that develop and maintain these tools.

Professor Shahramian's excellent video blog {\it The Signal Path} greatly informed this work. Dr. Dainis' videos on microbiology were also particularly useful.

Thanks to all the professors at York and beyond for returning my emails and checking their biofridges for phages.

Prof. Griffiths and Dr. Brovko at UGuelph, Dr. Anany at the GRDC for tips on the Beta-Glo assay

Professor Z-G Wang at Caltech for a note on electrostatics,

Profs David and Taylor (among others) at YorkU for being super nice, and H.P. and ice9 for tolerating ramblings.

Thanks to L.H. for wisdom regarding this project;

Special thanks to all the people - all equally capable of writing this paper if provided the opportunity - who circumstance led to work menial shifts at the component suppliers that remained open that were so critical. Hopfstader be damned; despite the name below the abstract, this paper is of course entirely the result of circumstance and luck.

R.I.P. to Dr. Robert Adair, whose work features prominently here, who died Sep. 2020.

\rule{\linewidth}{0.2pt}

The authors reluctantly acknowledge the Computational Electromagnetics Group at The University of Texas at Austin for going through the trouble to develop the AustinMan and AustinWoman human body models and then making them unavailable at https://sites.utexas.edu/austinmanaustinwomanmodels/, and also not replying to email.

It's okay, I'm not bitter.

\rule{\linewidth}{0.2pt}

The authors gratefully acknowledge the IT.IS foundation for the same, and also compiling the tissue property databases.

\rule{\linewidth}{0.2pt}

One Very Important Thought

{\Large {\verbatim <3}}

\end{document}