### 2020-11-15 documents/fluorescence.tex
Tube 3

\begin{document}

If time-domain filtering is sufficient,

Low background is required.

Xu's method \cite{Quantification2020}

Contrary to standard microscopy, you want as little excitation light to enter the objective as possible.

Nanodrop, using 280 nm absorbance.

the input edge is considered a clock; the minimum pulse width limits apply, but are not clearly specified in the datasheet. In this case (1/155 mhz) = 6 ns.