### 2020-12-15 documents/pseudo_soliton.tex

%!TeX root = pseudo_soliton
\documentclass[paper.tex]{subfiles}
\begin{document}

Precursor waves do not appear to be explicitly mentioned in IEEE standards, but are implicitly governed by the Fourier-spectrum clauses and more recent FCC ultrawideband regulations. Ultra-short wavepackets of a similar nature are widely used in MRI systems with a repetition rate up to 300 MHz. We can make some mechanistic claims.

\footnotetext{
	Sincere thanks go to the legendary Kurt Oughstun for taking the time to thoroughly answer some of our questions.
}

The prototypical Vircator (not varactor) pinched electron beam tubes

step recovery diodes

\begin{sidenote}
There are some extremely tricky subtleties in the simulation.  \cite{propagation1992}. In \cite{Comments1993}, Leubbers says FDTD and FFT both are better than Asympt, Ogh says no there's no problem, Asymptotic is really accurate. Frankly, I can't make heads or tails of any of these, not understanding the asymptotic method at all.
\end{sidenote}

The 2018 NATO Research Task Group 189 report \cite{treatyelectromagnetic} is an excellent source of data for these.

\begin{quote}
	Other basic bioeffects research [50], [52], [53], [66], [81], [83] has failed to provide support for Albanese’s multiple	theoretical postulates.
\end{quote}

The real killer

\footnotemark
\footnotetext{Honestly, how cool is this? Like, this unimportant factor (as simple as the rise time!) connects so deeply with Schrodinger's wave-packets - the very same functions that apply for probability distributions - and produces a result so striking - a complete change }

The "asymptotic method", and the fourier series method. Unfortunately, the fourier coefficients and phases provide little intuitive sense for how different input pulses propagate.

%The Somerfield precursor is high-frequency,

\end{document}