### 2020-12-16 documents/acknowledgement.tex
Writing notes

%!TeX root = acknowledgement
\documentclass[paper.tex]{subfiles}
\begin{document}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Acknowledgements}\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This paper is not novel enough to deserve this chapter, but those to be acknowledged are.

The authors of the original papers deserve the credit for this entire work; it is rare that such a monumental finding is served on a silver platter.

Chiefly among the enablers are of course my parents, Doris and Brian, without whom this work would have been done by someone else, and who also supported this work financially.

Yuriy S at Indiana, and Prof. Sarris at UofT.

The legend of dispersion himself, Professor Oughstun, for sending me some thoughts and his paper.

Thanks to (Dr. ) P. Barfuss for spending time reviewing an early draft.

This paper would have been absolutely impossible without the support of thousands of open-source software packages, ranging from monolithic state-supported tools like Paraview to poorly-compensated but equally invaluable scripts like zonca. A special thanks to the dedicated authors that develop and maintain these tools.

Professor Shahramian's excellent video blog {\it The Signal Path} greatly informed this work. Dr. Dainis' videos on microbiology were also particularly useful.

Thanks to all the professors at York and beyond for returning my emails and checking their biofridges for phages.

Sandra at the Resources Sharing Department at YorkU for spending an hour looking for the Gray paper.

Prof. Griffiths and Dr. Brovko at UGuelph, Dr. Anany at the GRDC for useful tips on the Beta-Glo assay.

Professor Z-G Wang at Caltech for a note on viral electrostatics,

Profs H. Kassiri, C. David and W. Taylor (among others) at YorkU for encouragement, and H.P. and ice9 for tolerating midnight ramblings.

Thanks to L.H. for wisdom regarding this project, and to my friend Eric and friends at SimCoLab.

Special thanks to all the people - all equally capable of writing this paper if provided the opportunity - who circumstance led to work menial shifts at the component suppliers that remained open that were so critical. Hopfstader be damned; despite the name below the abstract, this paper is of course entirely the result of circumstance and luck.

R.I.P. to Dr. Robert Adair and Dr. Lewis Franks, whose work features prominently here, died 2020 and 2019.

\rule{\linewidth}{0.2pt}

The authors gratefully acknowledge the IT.IS foundation

and also compiling the tissue property databases.

\rule{\linewidth}{0.2pt}

Thanks to Biotium Inc for support on GelGreen.

\rule{\linewidth}{0.2pt}

One Very Important Thought

{\Large {\verbatim <3}}

\end{document}