### 2021-01-18 documents/personal_device.tex

%!TeX root = personal_device
\documentclass[paper.tex]{subfiles}
\begin{document}

\begin{autem}
What follows is intended to be a casual, first-principles e. Be very careful with its conclusions.
\end{autem}

{\color{red} speculation \{ }

\clearpage
{\Large \it The plausibility of low-cost microwave masks}\\

With the most recent few generations of semiconductor processes, producing significant powers in the X-band has become inexpensive. Silicon transistors are limited in their output power and gain; Silicon-on-insulator topologies

This is perhaps most readily illustrated with the commercial HB100 radar module. This device implements a complete Doppler radar with two RF FETs and one dielectric resonator on a standard FR4 substrate. It operates at 10.6 GHz, yet are produced for under \$3.

Our crude experimentation has yielded about 10 mW with a single SiGe:C transistor, the Infineon BFP620 (\$ 0.34 apiece \@ 1000-of.). For the personal microwave sterilizer, given the threshold levels found by Yang, this corresponds to a cost per sterilized cubic centimeter of approx. \cite{BFP620H7764XTSA1}.

Trivially extending great work by \cite{Focusing} on passive reflectarrays, a 40-element active-antenna phased array, each element of which consists of a negative-resistance device such as the BFP620 transistor, a tuning varactor, a patch antenna, and a resonator, each contributing a pulsed tone of 500 nanoseconds length and 10 milliwatts of RF power, and a focal point spatial scan rate and pulse repetition frequency of 6 KHz. produces a field pattern which, when scanned, appears to be suitable to deny respiratory transmission in a 'face shield' configuration.

Such a field would comply with all standards by at least a factor of 20.

%figure from ipynb

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{muller_test_cropped.jpg}
	\caption{\\ How simple an active antenna element can be. This is based on the feedback-loop oscillator of \cite{SmallSize2008}, and uses a single CEL CE3520K3 GaAs FET (\$2.5) and one 100pF capacitor. It oscillated with extreme instability at about 5 GHz.}
\end{figure}

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{my_photo-17.jpg}
	\caption{\\ A wideband tunable 7 GHz VCO based on \cite{TripleTuned2008}. This uses an Infineon BFP620 SiGe:C transistor (\$0.3) and a network of four varactor diodes. Unfortunately, the Si varactors did not appear to have a sufficiently high Q to prevent mode-hopping, and the output spectrum reflects the ignorance of its designer; however, it appeared to be capable of producing the 10 dBm required above.}
\end{figure}

% a prototype oscillator

Injection locking rather than discrete phase shifters saves a number of components.

Such a device would cost about \$12 in silicon at prototype prices, provide half a month of battery life on two alkaline AAA cells,

We demonstrate this form factor because, superficially, there are fewer people than there are places. But is such a device really necessary or useful? Is it sufficiently superior to a fiber mask to justify spending energy developing it? It has the advantage of being self-cleaning; and hair, eyes; a hand brought near the face is immediately sterilized.

On the other hand, a personal device may present issues with participation and production volume.

Interference

If HFSS or ADS are not available, coupled SPICE-FDTD methods appear to be particularly simple and effective for these types of problems.

The advantage of the phased array is in the free-space power combining and the near-field focusing.

In our crude experiments, the loss tangent and poor impedance control of inexpensive substrates did not appear to pose a problem in itself; standard FR-4 or even PET substrate and screen-printed metal techniques seen in high-volume RFID tag production may be of some use.

However, the Q factor of tuning elements on such lossy materials appears to be too high to form useful oscillators. Air-dielectric coaxial, Barium titanate dielectric slab resonators, or various types of substrate-independent resonators may all be suitable for this purpose.

{\color{red} \} speculation }

\end{multicols}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\Large Interference}\\
\begin{multicols}{1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\Large Mass production}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Almost all components on the device can be replicated with fully vertically-integrated first-principles. Capacitors can be

If this 'electromagnetic mask' form is the ideal (not nearly),

The largest RFID plants can produce

a minimum of 5 GaAs or SiGe:C transistors will be required. Without SOI or, it does not appear that

As a lower bound, there are 1 million hospital beds in the U.S. [AHA 2018]; and as an upper bound obtaining global herd immunity would take 1.75 billion units.

It is difficult to determine the supply capacity for these semiconductor processes; information is not forthcoming from the manufacturers. Fermi estimates \footnote{GaAs MMIC market of \$2.2 Bn USD / random sample of MMIC prices = about 2 Bil devices / yr, 3e9 wifi connected devices produced each year,}

It is possible to use common Si-based devices at these frequencies, especially with second-harmonic techniques [Winch 1982]. However, obtaining the required gain and output power is not a trivial matter.

The techniques and equipment required to produce these devices are extremely complex; load-lock UHV,

Figures are not forthcoming,

Given the supply difficulties of comparatively simple materials such as Tyvek at pandemic scales, it is difficult to imagine that RF semiconductor production can be immediately re-tasked and scaled to this degree.

\paragraph{\textbf{Vacuum RF triode}}\

Especially combined with an integrated titanium sorption pump,

One concern is the large filament heater power, which prevents the use of low-cost button cells for power. Use of cold-cathode field-emitter arrays would alleviate this issue, but at the cost of complexity.

Small tungsten incandescent lights are available down to 0.05 watts. With a suitable high-efficiency cathode coating, a pulsed heater power of less than 0.02

\end{multicols}

The X-band is also minimally absorbed by air, allowing action in the far-field and scaling to square-kilometer areas with single installations.

\clearpage

\end{document}