### 2021-01-28 documents/pseudo_soliton.tex
dammit the precursor's at the wrong frequency

%!TeX root = pseudo_soliton
\documentclass[paper.tex]{subfiles}
\begin{document}

Precursor waves do not appear to be explicitly mentioned in IEEE standards, but are implicitly governed by the Fourier-spectrum clauses and more recent FCC ultrawideband regulations. Ultra-short wavepackets of a similar nature are widely used in MRI systems with a repetition rate up to 300 MHz. We can make some mechanistic claims.

\footnotetext{
	Sincere thanks go to the legendary Kurt Oughstun for taking the time to thoroughly answer some of our questions.
}

\footnotemark
\footnotetext{Honestly, how cool is this? Like, this unimportant factor (as simple as the rise time!) connects so deeply with Schrodinger's wave-packets - the very same functions that apply for probability distributions - and produces a result so striking - a complete change }

\end{document}