### 2021-01-28 documents/pulse_generator.tex

%!TeX root = pulse_generator
\documentclass[paper.tex]{subfiles}
\begin{document}

\end{document}