### 2021-02-08 documents/paper.tex

\documentclass[fleqn,10pt]{article}

\include{preamble}

\begin{document}

\subfile{title}

\begin{table}[h!]
	\centering
	\begin{tabular}{ |c|c|c|c| } 
		\hline
		col1 & col2 & col3 \\
		\hline
		& cell2 & cell3 \\ 
		& cell5 & cell6 \\ 
		& Negative & 5 \\ 
		\hline
	\end{tabular}

	\caption{Data from testing on bacteriophage. Data rounded to the nearest thousand photon counts. Not thoroughly characterized. Repeatability between test conditions   Background was no. 
	The flurometer printed a constant stream of the integration, all within the random error between capture windows $\sigma=0.5$ kcounts. }
\end{table}

\section{Figures which have nothing to do with our conclusion but are pretty nonetheless}

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{pulse_exposure_setup.JPG}
	\caption{\\ \textit{Chengxiang} pulse exposure jig. pulser seen off the left. Autosampler Pulse shaping plate not shown. Aluminum disc belongs to the scan-conversion digitizer, not used in this study.}
\end{figure}

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{eppenwolf_2.jpg}
	\caption{\\ \textit{Eppenwolf} misguided pulsed X-Band microwave spectrometer used in this study, pictured with the near-field cuvette installed.}
\end{figure}

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{bronch_9GHz_500W_2_transparent.png}
	\caption{\\ \textit{Eppenwolf} misguided pulsed X-Band microwave spectrometer used in this study, pictured with the near-field cuvette installed.}
\end{figure}

\begin{figure}[H]
	\makebox[\textwidth][c]{
		\includegraphics[width=\textwidth*2]{chunk_4_line_2_1.png}
	}%
	\caption{Chunk 4, it.is Founda}
\end{figure}

\clearpage

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{CPWG_sim_pretty_2}
	\caption{FDTD simulation geometry of the 0.3 mm coplanar waveguide exposure cell. Red object is the fluid channel. Visualized with Paraview.
		\ghfile{covidinator/electronics/simple_fdtd/run/microfluidic_coplanar.py}}
\end{figure}

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includesvg[width=\textwidth]{compChargePlot}
	\caption{
		\light{
			\\
			We've got a thing\\
			that's called\\
			{\it Radar Love}\\}
		We've got a wave\\
		in the air}
\end{figure}

\clearpage

\subfile{background}

\clearpage

%\subfile{safety}

\subfile{brillouin_pulse_math}

%\paragraph{}

%\subfile{acknowledgement}

\subfile{charge}

\subfile{viscosity}

\subfile{molecular_dynamics}

\section{Appendix on microwave design}

\subfile{supplemental}

\section{Appendix on microbiology}

\subfile{sic_susceptor}

\subfile{detection}

\subfile{fluorescence}

\subfile{lessons_learned}

\nocite{*}

%\printbibliography[]
%\printbibliography[title={All references}]

\end{document}