### 2021-02-08 documents/pulse_generator.tex

%!TeX root = pulse_generator
\documentclass[paper.tex]{subfiles}
\begin{document}

Impedance matching to appears to be a solved problem. \cite{Microchamber2011} provide a good review, and offer a design using a high-impedance exposure cell in parallel with a 50 ohm resistor, which offers excellent performance to at least 1 GHz. Attenuator for diagnostics are also built in.

Almost   Diagnostics of such pulses are difficult.

 Attenuators with such high peak powers are difficult to obtain.

\cite{Characterization2012}

\end{document}