### 2021-02-09 documents/pulse_generator.tex

%!TeX root = pulse_generator
\documentclass[paper.tex]{subfiles}
\begin{document}

Spark gaps

\cite{kHz1995} uses an interesting construction of a pressurized spark gap.

The ZTX417 parts are expensive (\$15 apiece, as of this writing).

[] mention triggering parameters

There seem to be two main designs:

drivers Pockels cells for Q switching.

Schaal patent  which can produce 70 ps edges using less expensive ZTX415 (although the all-important RF "second schematic", the ).

marx or marx-free

Almost   Diagnostics of such pulses are difficult.

Attenuators that are rated for such high peak powers are hard to find.

Oak \cite{fast1991} Figure 1 offers a circuit to evaluate avalanche behavior of transistors. Using this, the

Impedance matching to appears to be a solved problem. \cite{Microchamber2011} provide a good review, and offer a design using a high-impedance exposure cell in parallel with a 50 ohm resistor, which offers excellent performance to at least 1 GHz. Attenuator for diagnostics are also built in. This cell should be used rather than our design. \cite{Characterization2012}

\end{document}