### 2021-02-17 documents/acknowledgement.tex
Tersenating

%!TeX root = acknowledgement
\documentclass[paper.tex]{subfiles}
\begin{document}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Conflicts of interest

Funded in part by the author's company, SafeSump Inc., and Ideal Consulting Inc., whatever that's worth.

This paper is not novel enough to deserve this chapter, but those to be acknowledged are.

Special thanks to my parents, Doris and Brian, without whom this work would have been accomplished by someone else, and whose extraordinary support made this paper possible.

Sci-Hub - a very important thought

The authors of the original papers deserve the credit for this entire work; it is rare that such a monumental finding is served on a silver platter.

Yuriy S at Indiana, and Prof. Sarris at UofT. Sean Hum and [that other guy?] from UofT.

The legend of dispersion himself, Professor Oughstun, for sending me some thoughts and his paper.

Thanks to (Dr. ) P. Barfuss for spending time reviewing an early draft.

This paper would have been absolutely impossible without the support of thousands of open-source software packages, ranging from monolithic state-supported tools like Paraview to poorly-compensated but equally invaluable scripts like zonca. A special thanks to the dedicated authors that develop and maintain these tools.

Professor Shahramian's excellent video blog {\it The Signal Path} greatly informed this work. Dr. Dainis' videos on microbiology were also particularly useful.

Thanks to all the professors at York and beyond for returning my emails and checking their biofridges for phages.

Sandra and colleagues at the Resources Sharing Department at YorkU for spending an hour looking for 
the Gray paper.

Prof. Griffiths and Dr. Brovko at UGuelph, Dr. Anany at the GRDC for useful tips on the Beta-Glo assay.

Professor Z-G Wang at Caltech for a note on viral electrostatics,

Profs H. Kassiri, C. David and W. Taylor (among others) at YorkU for encouragement, and friends on twitter for tolerating midnight ramblings.

Thanks to L.H. for wisdom regarding this project, and to my friend Eric, and friends at SimCoLab.
%
%Special thanks to all the people - all equally capable of writing this paper if provided the opportunity - who circumstance led to work menial shifts at the component suppliers that remained open that were so critical. Hopfstader be damned; despite the name below the abstract, this paper is of course entirely the result of circumstance and luck.

R.I.P. to Dr. Robert Adair and Dr. Lewis Franks, whose work features prominently here, died 2020 and 2019.

Greenland for patience

James Hawthorne for Autoleveller

Chengxiang Li and the team of who appear to have spent a half-decade iterating the Marx avalanche generator we use in this paper. It is an excellent, very high performance design.

\rule{\linewidth}{0.2pt}

The authors gratefully acknowledge the IT.IS foundation

and also compiling the tissue property databases.

\rule{\linewidth}{0.2pt}

Thanks to Biotium Inc for support on GelGreen.

\rule{\linewidth}{0.2pt}

One Very Important Thought

{\Large 
	{
\begin{verbatim}
		<3 <3 <3
\end{verbatim}
}}

\end{document}