### 2021-02-17 documents/paper.tex

\documentclass[fleqn,10pt]{article}

\include{preamble}

\begin{document}

\subfile{title}

\section{Figures which have little to do with our conclusion but are pretty nonetheless}

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{pulse_exposure_setup.JPG}
	\caption{\\ \textit{Chengxiang} nanosecond pulse exposure jig. Pulser seen off the left. Pulse shaping plate not shown. Aluminum disc belongs to the scan-conversion digitizer, not used in this study.}
\end{figure}

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{eppenwolf_2.jpg}
	\caption{\\ \textit{Eppenwolf} misguided pulsed X-Band microwave spectrometer used in this study. (cuvette shown was replaced by the same )}
\end{figure}

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{bronch_9GHz_500W_2_transparent.png}
	\caption{Simulation of a (largely ineffective) bronchoscopic method.}
\end{figure}

%
%\begin{figure}[H]
%	\makebox[\textwidth][c]{
%		\includegraphics[width=\textwidth*2]{chunk_4_line_2_1.png}
%	}%
%	\caption{Chunk 4, it.is Founda}
%\end{figure}

%\clearpage

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{CPWG_sim_pretty_2}
	\caption{FDTD simulation geometry of the 0.3 mm coplanar waveguide exposure cell. Red object is the fluid channel. Visualized with Paraview.
		\ghfile{covidinator/electronics/simple_fdtd/run/microfluidic_coplanar.py}}
\end{figure}

\begin{figure}[H]
	%	\makebox[\textwidth][c]{
	\includesvg[width=\textwidth]{propagated_waveforms}
	\caption{Brillouin precursor propagation through tissue (reproduced from section 6).\\
	(a) Gaussian monopulse with 16 ps FWHM.\\
	(b) Optimized 10 GHz sawtooth wave. Note the slight high-frequency ripple.\\ This has a frequency of approximately 1.5 GHz.\\
	(c) Prototypical precursor 10 GHz sine burst.\\
	(d) Continuous 10 GHz sine tone.}

\end{figure}

%\begin{figure}[H]
%	\captionsetup{singlelinecheck = false, justification=justified}
%	\centering
%	\includesvg[width=\textwidth]{compChargePlot}
%	\caption{Charge distribution using Poisson Boltzmann}
%\end{figure}
%

\clearpage

\subfile{background}

\clearpage

%\subfile{safety}

\subfile{brillouin_pulse_math}

%\paragraph{}

%\subfile{acknowledgement}

\subfile{charge}

\subfile{viscosity}

\subfile{experiment_1}

\subfile{experiment_2}

\subfile{molecular_dynamics}

\section{Appendix on microwave design}

\subfile{supplemental}

\subfile{personal_device}

\section{Appendix on microbiology}

\subfile{sic_susceptor}

\subfile{detection}

\subfile{fluorescence}

\subfile{lessons_learned}

\subfile{acknowledgement}

\nocite{*}

% LC_ALL=C tr -dc '\0-\177' <references.bib >newfile && mv newfile references.bib

\printbibliography[title={All references}]

\end{document}