### 2021-02-17 documents/personal_device.tex

%!TeX root = personal_device
\documentclass[paper.tex]{subfiles}
\begin{document}
%
%\begin{autem}
%What follows is intended to be a casual, first-principles e. Be very careful with its conclusions.
%\end{autem}
%
%{\color{red} speculation \{ } 
%
%%\clearpage
{\Large \it The plausibility of low-cost microwave masks}\\

With the most recent few generations of semiconductor processes, producing significant powers in the X-band has become inexpensive. Silicon transistors are limited in their output power and gain; Silicon-on-insulator topologies

This is perhaps most readily illustrated with the commercial HB100 radar module. This device implements a complete Doppler radar with two RF FETs and one dielectric resonator on a standard FR4 substrate. It operates at 10.6 GHz, yet are produced for under \$3.

Our crude experimentation has yielded about 10 mW with a single SiGe:C transistor, the Infineon BFP620 (\$ 0.34 apiece \@ 1000-of.). For the personal microwave sterilizer, given the threshold levels found by Yang, this corresponds to a cost per sterilized cubic centimeter of approx. \cite{BFP620H7764XTSA1}.

Trivially extending great work by \cite{Focusing} on passive reflectarrays, a 40-element active-antenna phased array, each element of which consists of a negative-resistance device such as the BFP620 transistor, a tuning varactor, a patch antenna, and a resonator, each contributing a pulsed tone of 500 nanoseconds length and 10 milliwatts of RF power, and a focal point spatial scan rate and pulse repetition frequency of 6 KHz. produces a field pattern which, when scanned, appears to be suitable to deny respiratory transmission in a 'face shield' configuration.

Such a field would comply with all standards by at least a factor of 20.

%figure from ipynb

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{muller_test_cropped.jpg}
	\caption{\\ How simple an active antenna element can be. This is based on the feedback-loop oscillator of \cite{SmallSize2008}, and uses a single CEL CE3520K3 GaAs FET (\$2.5) and one 100pF capacitor. It oscillated with extreme instability at about 5 GHz.}
\end{figure}

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includegraphics[width=\textwidth]{my_photo-17.jpg}
	\caption{\\ A wideband tunable 7 GHz VCO based on \cite{TripleTuned2008}. This uses an Infineon BFP620 SiGe:C transistor (\$0.3) and a network of four varactor diodes. Unfortunately, the Si varactors did not appear to have a sufficiently high Q to prevent mode-hopping, and the output spectrum reflects the ignorance of its designer; however, it appeared to be capable of producing the 10 dBm required above.}
\end{figure}

\begin{figure}[H]
	\captionsetup{singlelinecheck = false, justification=justified}
	\centering
	\includesvg[width=\textwidth]{personal_device_E}
	\caption{Near-field focused phased array}
\end{figure}

% a prototype oscillator

Injection locking rather than discrete phase shifters saves a number of components.

Such a device would cost about \$12 in silicon at prototype prices, provide half a month of battery life on two alkaline AAA cells,

We demonstrate this form factor because, superficially, there are fewer people than there are places. But is such a device really necessary or useful? Is it sufficiently superior to a fiber mask to justify spending energy developing it? It has the advantage of being self-cleaning; and hair, eyes; a hand brought near the face is immediately sterilized.

On the other hand, a personal device may present issues with participation and production volume.

Interference

If HFSS or ADS are not available, coupled SPICE-FDTD methods appear to be particularly simple and effective for these types of problems.

The advantage of the phased array is in the free-space power combining and the near-field focusing.

In our crude experiments, the loss tangent and poor impedance control of inexpensive substrates did not appear to pose a problem in itself; standard FR-4 or even PET substrate and screen-printed metal techniques seen in high-volume RFID tag production may be of some use.

However, the Q factor of tuning elements on such lossy materials appears to be too high to form useful oscillators. Air-dielectric coaxial, Barium titanate dielectric slab resonators, or various types of substrate-independent resonators may all be suitable for this purpose.

{\color{red} \} speculation }

%\end{multicols}

%\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%{\Large Interference}\\
%\begin{multicols}{1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%{\Large Mass production}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%Almost all components on the device can be replicated with fully vertically-integrated first-principles. Capacitors can be
%
%If this 'electromagnetic mask' form is the ideal (not nearly), 
%
%The largest RFID plants can produce
%
%a minimum of 5 GaAs or SiGe:C transistors will be required. Without SOI or, it does not appear that 
%
%As a lower bound, there are 1 million hospital beds in the U.S. [AHA 2018]; and as an upper bound obtaining global herd immunity would take 1.75 billion units.
%
%It is difficult to determine the supply capacity for these semiconductor processes; information is not forthcoming from the manufacturers. Fermi estimates \footnote{GaAs MMIC market of \$2.2 Bn USD / random sample of MMIC prices = about 2 Bil devices / yr, 3e9 wifi connected devices produced each year,}
%
%
%
%It is possible to use common Si-based devices at these frequencies, especially with second-harmonic techniques [Winch 1982]. However, obtaining the required gain and output power is not a trivial matter. 
%
%The techniques and equipment required to produce these devices are extremely complex; load-lock UHV,  
%
%
%Figures are not forthcoming, 
%
%Given the supply difficulties of comparatively simple materials such as Tyvek at pandemic scales, it is difficult to imagine that RF semiconductor production can be immediately re-tasked and scaled to this degree. 
%
%\paragraph{\textbf{Vacuum RF triode}}\
%
%
%
%Especially combined with an integrated titanium sorption pump,
%
%One concern is the large filament heater power, which prevents the use of low-cost button cells for power. Use of cold-cathode field-emitter arrays would alleviate this issue, but at the cost of complexity.
%
%Small tungsten incandescent lights are available down to 0.05 watts. With a suitable high-efficiency cathode coating, a pulsed heater power of less than 0.02
%
%
%
%\end{multicols}
%
%
%The X-band is also minimally absorbed by air, allowing action in the far-field and scaling to square-kilometer areas with single installations.
%
%
%
%
%\clearpage
%

\end{document}