### 2021-02-22 documents/experiment_1_data.tex
More edits

%!TeX root = experiment_1_data
\documentclass[paper.tex]{subfiles}
\begin{document}

\begin{table}[h!]
	\centering
	\begin{tabular}{ |c|c|c| } 
		\hline
		Exposure & kilocounts per 10s integration & Note \# \\
		\hline
		Autoclaved phage (0.2 mL) & 30 & \\ 
		Autoclaved p. (50 uL) & 32 & Run 8\\ 
		\hline
		Phage culture 'blank' & 8 & \\ 
		Water & 1 & \\
		\hline
		\multicolumn{3}{|c|}{\ntilde 10 or 20 us  } \\
		\hline
		 & \\ 
		\hline
		\multicolumn{3}{|c|}{Sweep from 6 to 12 GHz, average approximately 150 V/m, 10 or 20 us CW pulses} \\
		\hline
		\multicolumn{3}{|c|}{\ntilde 0.05 mm, \ntilde } \\
		\hline
		\multicolumn{3}{|c|}{\ntilde 0.2 mm channel, 0.2 uL/s, various pulse counts and termination} \\
		\hline 
		 Blank    & 8 & Li test 1\\
		 +           & 8 & Li test 1\\
 		 +           & 7 & Li test 1\\
 		 +           & 7 & Li test 1\\
 		 + \& shaper & 7 & Li test 1\\
 		 \hline
		\multicolumn{3}{|c|}{\ntilde 0.05 mm channel, \ntilde 0.75 MV/m, 0.2 uL/s, various pulse counts and termination} \\
		\hline
		Sham & 5 & Smaller \\
		+ & 6 & Smaller \\ 
		+ & 6 & Smaller  \\ 
		\hline
		\multicolumn{3}{|c|}{\ntilde 0.05 mm channel, \ntilde 0.75 MV/m, 0.2 uL/s, basic buffer, but pH may not have been 10.6} \\
		\hline
		+ & 11 & Weakened \\
		+ (no term) & 11 & Weakened \\ 
		Sham & 12 & Weakened  \\ 
		\hline
	\end{tabular}
\end{table}

Data from testing on bacteriophage. Data rounded (0.5 and above up) to the nearest thousand photon counts per 10 seconds. %thoroughly characterized.

Background ('blank', in notes) readings measure the amount of DNA present in the phage culture tube before any treatment.

Repeatability between test conditions was not terrific. The source of the \ntilde 5000 count variation between some equivalent samples is not known. However, in no case did any tube approach the 22 to 30 kcounts of the autoclave.

Stock culture had titer $3\times 10^9$ T4r.

The fluor-meter printed a constant stream of results, all within the random error between capture windows $\sigma=0.5$ kcounts. Only after most of the data was taken was it realized how dumb this was.

Results aggregated over the course of several months.

In all cases, a thermal effect was implausible; the total energy in the pulse train precluded heating above the capsid melting temperature.\cite{Effects1951}

In an early iteration of the experiment, the firmware and data analysis scripts randomly generated blinded sham and autoclave assignments and automatically treated the samples accordingly. It also printed blinded templates for pipetting. This was done partly in the hopes that massively oversampling the microwave absorption spectrometer across the range of physical tolerances would highlight any resonances.\cite{first2000}

However, as spectrometry was abandoned, this non-deterministic property caused significant issues while troubleshooting the fluorescence detection system, and \st{expedience} \st{\textit{laziness}} prompted us to disable this system, violating best practices.

A thin layer of silicone conformal coating was applied to the 0.2 mm waveguide to avoid the copper ion artifact, but this could not be done on the 0.05 mm waveguide.

\end{document}