### 2021-03-01 documents/experiment_1_data.tex

%!TeX root = experiment_1_data
\documentclass[paper.tex]{subfiles}
\begin{document}

\subsection{Data from testing on bacteriophage T4}
\begin{table}[h!]
	\centering
	\begin{tabular}{ |c|c|c| } 
				\hline
		\multicolumn{3}{|c|}{\textbf{Calibration trials} } \\
		\hline
		\hline
		Exposure & kilocounts per 10s integration & Note \# \\
		\hline
		Autoclaved phage (0.2 mL) & 30 & \\ 
		Autoclaved p. (50 uL) & 32 & Run 8\\ 
		\hline
		Phage culture 'blank' & 8 & \\ 
		Water & 1 & \\

		\hline
\multicolumn{3}{|c|}{\textbf{Pulsed CW trials}} \\
\hline

		\hline
		\multicolumn{3}{|c|}{\textbf{NF capillary}, 6 to 12 GHz, 10 or 20 us CW pulses} \\
		\hline
		Sham   & 12 & Run 8 t4,5...\\
		+ & 11 & Run 8 t4,5...\\
		+  & 12 & Run 8 t4,5...\\
		+  & 12 & Run 8 t4,5...\\
		\hline
		\multicolumn{3}{|c|}{\textbf{NF capillary}, 6 to 12 GHz, phage degraded with basic buffer} \\
		\hline
		Blank (0 min) & 14 & Run 9\\
		Blank (\ntilde 15 min) & 17 & Run 9\\
		Sham (\ntilde 15 min) & 5 & Run 9\\
		Blank (\ntilde 4 hr) & 24 & Run 9\\
		\hline
		+  & 14 & Run 11\\
		Sham & 17 & Run 11\\
		\hline
		\multicolumn{3}{|c|}{\textbf{NF capillary}, 6 to 12 GHz, 10 or 20 us CW pulses} \\
		\hline
		Sham & 12 & Run 12\\
		+    & 10 & Run 12\\
		\hline
		\multicolumn{3}{|c|}{\textbf{0.2 mm channel}, 6 to 12 GHz, 10 or 20 us CW pulses} \\
		\hline
		 Sham & 12 & CPWG micro-\\
 		 +    & 10 & CPWG micro-\\

		\hline
		\multicolumn{3}{|c|}{\textbf{Nanosecond pulsed trials} } \\
		\hline

 		\hline
		\multicolumn{3}{|c|}{\ntilde 0.2 mm channel, 0.2 uL/s, various pulse counts and termination} \\
		\hline 
		 Blank    & 8 & Li test 1\\
		 +           & 8 & Li test 1\\
 		 +           & 7 & Li test 1\\
 		 +           & 7 & Li test 1\\
 		 + \& shaper & 7 & Li test 1\\
 		 \hline

		\multicolumn{3}{|c|}{\ntilde 0.05 mm channel, 0.2 uL/s, various pulse counts and termination} \\
		\hline
		Sham & 5 & Smaller \\
		+ & 6 & Smaller \\ 
		+ & 6 & Smaller  \\ 
		\hline

		\multicolumn{3}{|c|}{\ntilde 0.05 mm channel, 0.2 uL/s, basic buffer, tainted as pH may not have been 10.6} \\
		\hline
		Sham & 12 & Weakened \\ 
		Blank & 11 & Weakened  \\ 
		Blank & 12 & Weakened  \\ 
		Blank & 9 & Weakened  \\ 
		+ & 11 & Weakened \\
		+ (no term) & 11 & Weakened \\ 
		+ & 9 & Weakened \\ 
		\hline
	\end{tabular}
\end{table}

Original data for this table can be inspected at \ghfile{firmware/eppenwolf/pulse_1.pnw}.

See notes on data collection in main body.

\end{document}