### 2021-03-04 documents/experiment_1_data.tex
Fields put in data table

%!TeX root = experiment_1_data
\documentclass[paper.tex]{subfiles}
\begin{document}

\subsection{Data from testing on bacteriophage T4}
\begin{table}[h!]
	\centering
	\begin{tabular}{ |c|c|c|c| } 
				\hline
		\multicolumn{3}{|c|}{\textbf{Calibration trials} } \\
		\hline
		\hline
		Exposure & kilocounts per 10s integration & Note \# \\
		\hline
		Autoclaved phage (0.2 mL) & 30 & \\ 
		Autoclaved p. (50 uL) & 32 & Run 8\\ 
		\hline
		Phage culture 'blank' & 8 & \\ 
		Water & 1 & \\

		\hline
\multicolumn{3}{|c|}{\textbf{Pulsed CW trials}} \\
\hline

		\hline
		\multicolumn{3}{|c|}{\textbf{NF capillary}, 6 to 12 GHz, 10 or 20 us CW pulses 84.8 V/m pessimistic to 339 V/m optimistic} \\
		\hline
		Sham   & 12 & Run 8 t4,5...\\
		+ & 11 & Run 8 t4,5...\\
		+  & 12 & Run 8 t4,5...\\
		+  & 12 & Run 8 t4,5...\\
		\hline
		\multicolumn{3}{|c|}{\textbf{NF capillary}, 6 to 12 GHz, phage degraded with basic buffer, 84.8 V/m pessimistic to 339 V/m optimistic} \\
		\hline
		Blank (0 min) & 14 & Run 9 \\
		Blank (\ntilde 15 min) & 17 & Run 9\\
		Sham (\ntilde 15 min) & 5 & Run 9\\
		Blank (\ntilde 4 hr) & 24 & Run 9\\
		\hline
		+  & 14 & Run 11\\
		Sham & 17 & Run 11\\
		\hline
		\multicolumn{3}{|c|}{\textbf{NF capillary}, 6 to 12 GHz, 10 or 20 us CW pulses, 84.8 V/m pessimistic to 339 V/m optimistic} \\
		\hline
		Sham & 12 & Run 12\\
		+    & 10 & Run 12\\
		\hline
		\multicolumn{3}{|c|}{\textbf{0.2 mm channel}, 6 to 12 GHz, 10 or 20 us CW pulses, 200 V/m pes. and 10.6 kV/m opt.} \\
		\hline
		 Sham & 12 & CPWG micro-\\
 		 +    & 10 & CPWG micro-\\

		\hline
		\multicolumn{3}{|c|}{\textbf{Nanosecond pulsed trials} } \\
		\hline

 		\hline
		\multicolumn{3}{|c|}{\ntilde 0.2 mm channel, various pulse counts and termination, 0.155 MV/m to 5.6 MV/m} \\
		\hline 
		 Blank    & 8 & Li test 1\\
		 +           & 8 & Li test 1\\
 		 +           & 7 & Li test 1\\
 		 +           & 7 & Li test 1\\
 		 + \& shaper & 7 & Li test 1\\
 		 \hline

		\multicolumn{3}{|c|}{\ntilde 0.03 mm channel, 0.2 uL/s, various pulse counts and termination, 1 MV/m to 38 MV/m} \\
		\hline
		Sham & 5 & Smaller \\
		+ & 6 & Smaller \\ 
		+ & 6 & Smaller  \\ 
		\hline

		\multicolumn{3}{|c|}{\ntilde 0.03 mm channel, basic buffer, tainted as pH may not have been 10.6, 1 MV/m to 38 MV/m} \\
		\hline
		Sham & 12 & Weakened \\ 
		Blank & 11 & Weakened  \\ 
		Blank & 12 & Weakened  \\ 
		Blank & 9 & Weakened  \\ 
		+ & 11 & Weakened \\
		+ (no term) & 11 & Weakened \\ 
		+ & 9 & Weakened \\ 
		\hline
	\end{tabular}
\end{table}

Original notes for this table can be inspected at \ghfile{firmware/eppenwolf/pulse_1.pnw}. Data rounded to the nearest thousand fluorescence photon counts per 10 seconds. Larger numbers are presumed to correlate with greater capsid damage.

See notes on experiment design and data collection in main body.

\end{document}