### 2021-03-05 documents/experiment_1_data.tex
More work on writing, need to add pulse counts

%!TeX root = experiment_1_data
\documentclass[paper.tex]{subfiles}
\begin{document}

\subsection{Data from testing on bacteriophage T4}

\begin{table}[h!]
	\centering
	\begin{tabular}{ |c|c|c|c| } 
				\hline
		\multicolumn{3}{|c|}{\textbf{Calibration trials} } \\
		\hline
		\hline
		Exposure & kilocounts per 10s integration & Note \# \\
		\hline
		Autoclaved phage (0.2 mL) & 30 & \\ 
		Autoclaved p. (50 uL) & 32 & Run 8\\ 
		\hline
		Phage culture 'blank' & 8 & \\ 
		Water & 1 & \\

		\hline
\multicolumn{3}{|c|}{\textbf{Pulsed CW trials}} \\
\hline

		\hline
		\multicolumn{3}{|c|}{\textbf{NF capillary}, 6 to 12 GHz, 10 or 20 us CW pulses, 84.8 V/m to 339 V/m} \\
		\hline
		Sham   & 12 & Run 8 t4,5...\\
		+ & 11 & Run 8 t4,5...\\
		+  & 12 & Run 8 t4,5...\\
		+  & 12 & Run 8 t4,5...\\
		\hline
		\multicolumn{3}{|c|}{\textbf{NF capillary}, 6 to 12 GHz, phage degraded with basic buffer, 84.8 V/m to 339 V/m} \\
		\hline
		Blank (0 min) & 14 & Run 9 \\
		Blank (\ntilde 15 min) & 17 & Run 9\\
		Sham (\ntilde 15 min) & 5 & Run 9\\
		Blank (\ntilde 4 hr) & 24 & Run 9\\
		\hline
		+  & 14 & Run 11\\
		Sham & 17 & Run 11\\
		\hline
		\multicolumn{3}{|c|}{\textbf{NF capillary}, 6 to 12 GHz, 10 or 20 us CW pulses, 84.8 V/m to 339 V/m} \\
		\hline
		Sham & 12 & Run 12\\
		+    & 10 & Run 12\\
		\hline
		\multicolumn{3}{|c|}{\textbf{0.2 mm channel}, 6 to 12 GHz, 10 or 20 us CW pulses, 200 V/m pes. and 10.6 kV/m opt.} \\
		\hline
		 Sham & 12 & CPWG micro-\\
 		 +    & 10 & CPWG micro-\\

		\hline
		\multicolumn{3}{|c|}{\textbf{Nanosecond pulsed trials} } \\
		\hline

 		\hline
		\multicolumn{3}{|c|}{\ntilde 0.2 mm channel, various pulse counts and termination, 0.155 MV/m to 5.6 MV/m} \\
		\hline 
		 Blank    & 8 & Li test 1\\
		 +           & 8 & Li test 1\\
 		 +           & 7 & Li test 1\\
 		 +           & 7 & Li test 1\\
 		 + \& shaper & 7 & Li test 1\\
 		 \hline

		\multicolumn{3}{|c|}{\ntilde 0.03 mm channel, 0.2 uL/s, various pulse counts and termination, 1 MV/m to 38 MV/m} \\
		\hline
		Sham & 5 & Smaller \\
		+ & 6 & Smaller \\ 
		+ & 6 & Smaller  \\ 
		\hline

		\multicolumn{3}{|c|}{\ntilde 0.03 mm channel, basic buffer, tainted as pH may not have been 10.6, 1 MV/m to 38 MV/m} \\
		\hline
		Sham & 12 & Weakened \\  
		Blank & 11 & Weakened  \\ 
		Blank & 12 & Weakened  \\ 
		Blank & 9 & Weakened  \\ 
		+ & 11 & Weakened \\
		+ (no term) & 11 & Weakened \\ 
		+ & 9 & Weakened \\ 
		\hline
	\end{tabular}
\end{table}
%how many counts?

Note the width of the electric field error bars. These are not meaningful random or systematic intervals; while we have tried to do our best with the instruments available, we cannot confidently assert that the field was within this range.

Original notes for this table can be inspected at \ghfile{firmware/eppenwolf/pulse_1.pnw}. Data rounded to the nearest thousand fluorescence photon counts per 10 seconds. Larger numbers are presumed to correlate with greater capsid damage.

See notes on experiment design and data collection in main body.

\end{document}