### 2021-03-21 documents/EPA_paper.tex
Pure electroporation

%!TeX root = EPA_paper
\documentclass[paper.tex]{subfiles}
\begin{document}

\section{Gray literature}

There is an interesting citation in the Gray literature \cite{Index1981}, particularly because it seems to have progressed to the clinic:

FELLOWES, O. N. GRAY, O. S. SANDERS, M. SELECTIVE EFFECT OF EM ENERGY ON VIRUSES 
NEW YORK ACADEMY OF SCIENCES SYMPOSIUM 'EFFECT OF CONTROLLED EM ENERGY ON BIOLOGICAL SYSTEMS', 4 NOV. 1970

SCHLUTZ, C. A. GRAY, O. S. SANDERS, M. FELLOWES, O. N. 
THE EM CONTROLLED EFFECT OF ENERGY ON VIRUSES IN HUMAN BLOOD NEW YORK ACADEMY OF SCIENCES SYMPOSIUM 'EFFECT OF CONTROLLED EM ENERGY ON BIOLOGICAL SYSTEMS', 4 NOV. 1970

I can find no reference to such a conference in Annals of the NYAS, but it may be recorded in Transactions. It is a shame to see good work fade into the mist of time, especially since research in this field attained an acceptable level of quality even in the 1940s. O. N. Fellowes was of the Plum Island Animal Disease Laboratory. Oscar S. Gray was of Gray Industries Inc, Florida.

From Mr. Gray's U.S. patents, however, we are not given any reason to believe that this effect was non-thermal in nature.

Thanks to [] at york, and [] on the UWaterloo Discord.
\end{document}